Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-07-03 draftfinalreport_v4.tex
Changed the % naive and all subsequent changes
Many many changes, some minor, some substantial, due to the latest MoH
vaccine coverage rates provided.

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand, 2013} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 11\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve to measles, it was estimated that approximately 28\% would require additional vaccination to ensure measles does not persist, leading to an additional 128,034 effective vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially immunised population) values often include or exceed one, including 2013--2014 outbreaks that were ongoing at the time of analysis. As analysed from data until 12 June 2014, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of the first 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost of those measles cases that attend hospital is estimated to be \$1,877 per attending case. Approximately 17\% of measles cases attend hospital.
\item The mean public health service cost per case (not including hospitalisation costs) is \$1,765.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccination-related costs required to exceed approximately \$5,365 per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs.
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 28\% of the currently 11\% na\"{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@

<<echo=FALSE,include=FALSE,results=hide>>=
rm(list=ls())
#library(adehabitat)
library(classInt)
library(ggplot2)   # plotting library
library(grid)
library(gridExtra)
library(ggmap)
library(Hmisc)
library(maps)       # Provides functions that let us plot the maps
library(mapdata)    # Contains the hi-resolution points that mark out the countries.
library(maptools)
library(mapproj)
library(plyr)
library(pscl)
library(RColorBrewer)
library(rworldmap)
library(reshape)
library(reshape2)
library(rgeos)
library(rgdal)
library(raster)
library(sp)
library(utils)
library(xtable)
## get data
# set wd
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5), rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))

naive<-round(pop-(pop*impop))

## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
colnames(naive)<-c("AgeInYears","Naive","Cases")

dose1=as.factor(round(vac$Dose1Mths/12,0));
dose1<-summary(dose1)
dose1<-as.data.frame(dose1)
dose1$AgeInYears<-rownames(dose1)
dose2=as.factor(round(na.omit(vac$Dose2Mths/12,0)));
dose2<-summary(dose2)
dose2<-as.data.frame(dose2)
dose2$AgeInYears<-rownames(dose2)

## need to merge vaccination data with data
## to get years the cases were from
datav<-merge(vac,data,by="CaseCode",all=T)
##
datav$Dose1Mths[which(is.na(datav$Dose1Mths))] <- 9999
datav$Dose2Mths[which(is.na(datav$Dose2Mths))] <- 9999
#########################
datav$RptYear<-as.factor(datav$RptYear)
datav$SurvWeek<-as.factor(datav$SurvWeek)
datav$NZDep01<-as.factor(datav$NZDep01)
datav$NZDep06<-as.factor(datav$NZDep06)
datav$NZDep13<-as.factor(datav$NZDep13)
datav$Dose1Mths<-as.factor(datav$Dose1Mths)
datav$Dose2Mths<-as.factor(datav$Dose2Mths)

#data$AgeInYears<-as.factor(data$AgeInYears)
datav$EthnicityPrioritised<-as.factor(datav$EthnicityPrioritised)

timev<-aggregate( cbind( DiseaseName) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear+ Dose1Mths +Dose2Mths, data = datav , FUN=sum,na.rm=F,na.action=na.pass)

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

testv$VC<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,0,
ifelse(testv$Dose1Mths >= 0 & is.na(testv$Dose2Mths) == T,1,2))
AgeVac<-table(testv$VC,testv$AgeInYears)

row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)
AgeInYears<-(as.numeric(rownames(AgeVac)))
AgeVac<-cbind(AgeVac,AgeInYears)       
AgeInYears<-(as.numeric(rownames(pop)))
pop<-cbind(pop,AgeInYears)
colnames(pop)<-c("Population","AgeInYears")
AgeV<-merge(pop,AgeVac,by="AgeInYears",all=T)

## regression analyses - measles
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear, data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MELAA","Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth
test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

# want to set NZ Deprivation to the appropriate year

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))

testt$NZDep<-as.factor(testt$NZDep-1)

testtable<-aggregate( cbind( DiseaseName) ~ NZDep+AgeInYears+EthnicityPrioritised + RptYear,  data = testt , FUN=sum)
dim(testtable)
colnames(testtable)<-c("NZDep","Age","Ethnicity","Year","Cases")
head(testtable)

tnzd<-revalue(testtable$NZDep, c("1"="1-3","2"="1-3","3"="1-3","4"="4-7",
                                 "5"="4-7","6"="4-7","7"="4-7","8"="8-10",
                                 "9"="8-10","10"="8-10"));

testtable$NZDep<-tnzd
# head(testtable)
tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]

# latex(ttyr, file="",table.env=FALSE,rowname=NULL)

### denominator data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MELAA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth
head(test)
test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

# want to set NZ Deprivation to the appropriate year

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
head(testt)
tail(testt)
str(testt)
testt$NZDep<-as.factor(testt$NZDep-1)

testtable<-aggregate( cbind( DiseaseName) #+ as.numeric(SurvWeek))
~ NZDep+AgeInYears+EthnicityPrioritised + RptYear, 
data = testt , FUN=sum)
dim(testtable)
colnames(testtable)<-c("NZDep","Age","Ethnicity","Year","Cases")
head(testtable)

#tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
#"5"="1-5","6"="6-10","7"="6-10","8"="6-10",
#"9"="6-10","10"="6-10"));

tnzd<-revalue(testtable$NZDep, c("1"="1-3","2"="1-3","3"="1-3","4"="4-7",
                                 "5"="4-7","6"="4-7","7"="4-7","8"="8-10",
                                 "9"="8-10","10"="8-10"));

testtable$NZDep<-tnzd
# head(testtable)
tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
denom<-read.csv("NZDep2006Denominators.csv",header=T)
head(denom)
denom<-denom[,-c(9)]
summary(denom)
dim(denom)
denom<-denom[,1:18]
row.has.na <- apply(denom, 1, function(x){any(is.na(x))})
sum(row.has.na)
denomt <- denom[!row.has.na,]
head(denomt)
denomt<-denomt[denomt$Sex_code==99,]
denomt$Eth_Level<-as.factor(denomt$Eth_Level)
denomt<-denomt[denomt$Eth_Level %in% c("24","21","66",'22','77'),]
dm<-cbind(denomt[,c(3,6,9:18)])
head(dm)
dm<-dm[dm$Age_Label %in% c("A: 0- 5 yrs", "A: 6-17 yrs","A:18-24 yrs","A:25-64 yrs","A:65+ yrs"),]
dage<-revalue(dm$Age_Label, c("A: 0- 5 yrs"="<6", "A: 6-17 yrs"="6-17","A:18-24 yrs"="18-24","A:25-64 yrs"="25+","A:65+ yrs"="25+"));
dm$Age_Label<-dage
colnames(dm)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")
head(dm)
datadm<-dm[dm$Age=="<6",]
dataA<-(rbind(datadm,datadm)/2)
dataA<-round(dataA[,3:12])
dataA<-cbind(c(rep("0-2",5),rep("3-5",5)),rep(datadm$Ethnicity[1:5],2),dataA)
colnames(dataA)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")
ddm<-dm[!dm$Age=="<6",]
dm<-rbind(dataA,ddm)
dm$Ethnicity <- factor(dm$Ethnicity)
deth<-revalue(dm$Ethnicity, c("Asian (Prioritised)"="Asian","European (NZ European and Other European)"="European",
"Maori (Prioritised)"="Maori","MELAA"="MELAA","Pacific People (Prioritised)"="Pacific"));
dm$Ethnicity<-deth
head(dm)

## here's the additional fudge the MoH asked for.
## add 5 year olds to next class
dm[11:15,3:12]<-dm[11:15,3:12]+round(dm[6:10,3:12]/3,0)
dm[6:10,3:12]<-round(dm[6:10,3:12]*2/3,0)
## add 2 year olds to next class
dm[6:10,3:12]<-dm[6:10,3:12]+round(dm[1:5,3:12]/3,0)
dm[1:5,3:12]<-round(dm[1:5,3:12]*2/3,0)
## 
# rename
dm$Age<-revalue(dm$Age, c("0-2"="0-1", "3-5"="2-4","6-17"="5-17"));

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
popn
#pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
#"Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
#"Dep9"="6-10","Dep10"="6-10"));

pnzd<-revalue(popn$NZDep, c("Dep1"="1-3","Dep2"="1-3","Dep3"="1-3","Dep4"="4-7",
                            "Dep5"="4-7","Dep6"="4-7","Dep7"="4-7","Dep8"="8-10",
                            "Dep9"="8-10","Dep10"="8-10"));

popn$NZDep<-pnzd
head(popn)
tp<-aggregate( cbind(Popn) ~ NZDep+Age+Ethnicity, 
data = popn , FUN=sum)

tp$merge <- paste(tp$NZDep, tp$Age, tp$Ethnicity)
ttyr$merge <- paste(ttyr$NZDep, ttyr$Age, ttyr$Ethnicity)
ttyr <- ttyr[ttyr$Ethnicity!="None",]

tp$cases <- 0
cases <- matrix(0, length(tp$merge),1)
rownames(cases) <- tp$merge
cases[ttyr$merge,] <- ttyr$Cases
tp$cases <- cases
PerCapita<-(tp$cases/tp$Popn)
PerCap<-cbind(tp[,c(1:4,6)],round(PerCapita*10000,4))
colnames(PerCap)<-c("NZDep","Age","Ethnicity","Population","Cases","Per capita")

tp$perCap<-with(tp,tp$cases/tp$Popn*10000)
tpsub<-tp[!(tp$Ethnicity=="MELAA"),]
tpsub$ordert <- factor(tpsub$merge, as.character(tpsub$merge))

tpsub1<-tpsub[ order(tpsub$Age), ]
tpsub1$ordert <- factor(tpsub1$merge, as.character(tpsub1$merge))

#hist(tp$cases,xlab="Cases",main='Histogram of cases per category',col='grey',breaks=20)
tp$Ethnicity<- relevel(tp$Ethnicity, "European")
model<-glm(cases~Age*Ethnicity*NZDep+offset(log(Popn)),data=tp,family="quasipoisson")
summary(model)

model1<-update(model,~.-Age:Ethnicity:NZDep)
summary(model1)
anova(model1,test="F") # F test, not Chisq, because dispersion estimated by moments

model2<-update(model1,~.-Age:NZDep)
summary(model2)
anova(model2,test="F")

## use tpsub
#hist(tpsub$cases,xlab="Cases",main='Histogram of cases per category',col="darkgrey",breaks=20)
tpsub$Ethnicity<- relevel(tpsub$Ethnicity, "European")

model<-glm(cases~Age*Ethnicity*NZDep+offset(log(Popn)),data=tpsub,family="quasipoisson")
summary(model)

model1<-update(model,~.-Age:Ethnicity:NZDep)
summary(model1)
anova(model1,test="F")

#model2<-update(model1,~.-Ethnicity:NZDep)
#summary(model2)
#anova(model2,test="F")
model2<-update(model1,~.-Age:NZDep)
summary(model2)
anova(model2,test="F")

model3<-update(model2,~.-Ethnicity:NZDep)
summary(model3)
anova(model3,test="F")
anovap<-anova(model3,test="F")

res<-predict(model3)
#res<-predict(model2)

cor(exp(res),tpsub$cases)
cor.test(exp(res),tpsub$cases)

resan<-xtable(anova(model3, test="F")[-1,6, drop=FALSE])
xtb<-xtable(summary(model3)$coefficients[,-3])

AgeVac<-table(testv$VC,testv$AgeInYears)
row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)

DOB<-as.numeric(levels(testv$RptYear))[testv$RptYear]-testv$AgeInYears
#hist(DOB,breaks=100,col="grey",main="2007- 2014 measles cases",xlab="Date of Birth")

DOBVac<-table(testv$VC,DOB)
row.names(DOBVac)<-c("Unvaccinated","Dose1","Dose2")
DOBVac<-t(DOBVac)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
#naive[is.na(naive)] <- 0
colnames(naive)<-c("AgeInYears","Naive","Cases")
##
###

# proportions of the naive population the would be covered if vaccinated by age groups now
propvCohort<-c(
round((sum(naive$Naive[3:6])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:12])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:17])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:19])/sum(naive$Naive)),2),
round((sum(naive$Naive[12:17])/sum(naive$Naive)),2),
round((sum(naive$Naive[12:19])/sum(naive$Naive)),2),
round((sum(naive$Naive[20:26])/sum(naive$Naive)),2),
round((sum(naive$Naive[20:36])/sum(naive$Naive)),2),
round((sum(naive$Naive[20:47])/sum(naive$Naive)),2),
round((sum(naive$Naive[1:30])/sum(naive$Naive)),2),
round((sum(naive$Naive[1:45])/sum(naive$Naive)),2),
round((sum(naive$Naive[8:100])/sum(naive$Naive)),2))
yr_range<-c("2-5","5-11","5-16","5-18","11-16","11-18","19-25","19-30","19-46","0-29","0-44",">8")
propvCohort<-cbind(yr_range,propvCohort)
propvCohort<-as.data.frame(propvCohort)
write.table(propvCohort,"vaccine_proportions_Cohort.txt",sep=",",row.names=T,col.names=F)
propsv = NULL
for (i in 1:20 ) {
  propsv[i] = round((sum(naive$Naive[i:13])/sum(naive$Naive)),2)
}
propsvAge = NULL
for (i in 1:20 ) {
  propsvAge[i]=paste(c(min(i:20)-1, max(i:20)-1),collapse="-")
}

table120<-as.data.frame(propsv,propsvAge)
table120<-cbind(Age=rownames(table120),table120)
rownames(table120)<-1:20
colnames(table120)<-c("Age","Proportion")

propsvrev = NULL
for (i in 20:1 ) {
  propsvrev[i] = round((sum(naive$Naive[i:1])/sum(naive$Naive)),2)
}
propsvrevAge = NULL
for (i in 20:1 ) {
  propsvrevAge[i]=paste(c(min(i:1)-1, max(i:1)-1),collapse="-")
}

table201<-as.data.frame(propsvrev,propsvrevAge)
table201<-cbind(Age=rownames(table201),table201)
rownames(table201)<-1:20
colnames(table201)<-c("Age","Proportion")

propsv11 = NULL
for (i in 1:11 ) {
  propsv11[i] = round((sum(naive$Naive[i:11])/sum(naive$Naive)),2)
}
propsv11Age = NULL
for (i in 1:11 ) {
  propsv11Age[i]=paste(c(min(i:11)-1, max(i:11)-1),collapse="-")
}

table11<-as.data.frame(propsv11,propsv11Age)
table11<-cbind(Age=rownames(table11),table11)
rownames(table11)<-1:11
colnames(table11)<-c("Age","Proportion")

# write.table(table11,"vaccine_proportions_11yrOlds.txt",sep=",",row.names=F)
# write.table(table120,"vaccine_proportions_1_20yrOlds.txt",sep=",",row.names=F)
# write.table(table201,"vaccine_proportions_20_1yrOlds.txt",sep=",",row.names=F)

###
popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
         rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))

totdpdf<-read.csv("2013mapdata_tot.csv",header=T)
# output top countries
colnames(totdpdf)[1]<-c("country")
toptot <- totdpdf[order(-totdpdf$immigration),] 
immigration<-toptot[1:30,c(1,4)]
topincidence <- totdpdf[order(-totdpdf$incidence),] 
incidence<-topincidence[1:30,c(1,2)]
incidence[,2]<-round(incidence[,2],0)
topvaccine <- totdpdf[order(totdpdf$cover),] 
vaccinecover<-topvaccine[1:30,c(1,3)]
toprisk <- totdpdf[order(-totdpdf$risk),] 
risk<-(toprisk[1:30,c(1,5)])
risk[,2]<-round(risk[,2],0)
toprisk2<-(toprisk[1:10,c(1,2,3,4)])
toprisk2[,2]<-round(toprisk2[,2],0)

nzdpdf<-read.csv("2013mapdata_nzers.csv",header=T)
# output top countries
colnames(nzdpdf)[1]<-c("country")
topnz <- nzdpdf[order(-nzdpdf$immigration),] 
nz<-topnz[1:30,c(1,4)]
topnzrisk <- nzdpdf[order(-nzdpdf$risk),] 
nzrisk<-(topnzrisk[1:30,c(1,5)])
nzrisk[,2]<-round(nzrisk[,2],0)

imdpdf<-read.csv("2013mapdata_immigration.csv",header=T)
# output top countries
colnames(imdpdf)[1]<-c("country")
topim <- imdpdf[order(-imdpdf$immigration),] 
im<-topim[1:30,c(1,4)]
topimrisk <- imdpdf[order(-imdpdf$risk),] 
imrisk<-(topimrisk[1:30,c(1,5)])
imrisk[,2]<-round(imrisk[,2],0)
# haven't had time to sort out which packages I need

change<-read.csv("change.csv",header=T)
costtable1<-read.csv("costtable1.csv",header=T)
costtable2<-read.csv("costtable2v2.csv",header=T)
costtable3<-read.csv("costtable3.csv",header=T)
costtable4<-read.csv("costtable4v2.csv",header=T)
costtable5<-read.csv("costtable5v2.csv",header=T)
costtable20<-read.csv("cost_benefit_20.csv",header=T)
costtable50<-read.csv("cost_benefit_50.csv",header=T)

vacp<-read.csv("vacc_predictions.csv",header=T)
hosp<-read.csv("hospital.csv",header=T)

t12006<-read.csv("MMR1summary2006.txt",header=T,sep="")
t22006<-read.csv("MMR2summary2006.txt",header=T,sep="")
t12007<-read.csv("MMR1summary2007.txt",header=T,sep="")
t22007<-read.csv("MMR2summary2007.txt",header=T,sep="")
t12008<-read.csv("MMR1summary2008.txt",header=T,sep="")
t22008<-read.csv("MMR2summary2008.txt",header=T,sep="")
t12009<-read.csv("MMR1summary2009.txt",header=T,sep="")
t22009<-read.csv("MMR2summary2009.txt",header=T,sep="")
t12010<-read.csv("MMR1summary2010.txt",header=T,sep="")
t22010<-read.csv("MMR2summary2010.txt",header=T,sep="")
t12011<-read.csv("MMR1summary2011.txt",header=T,sep="")
t22011<-read.csv("MMR2summary2011.txt",header=T,sep="")
t12012<-read.csv("MMR1summary2012.txt",header=T,sep="")
t22012<-read.csv("MMR2summary2012.txt",header=T,sep="")
t12013<-read.csv("MMR1summary2013.txt",header=T,sep="")
t22013<-read.csv("MMR2summary2013.txt",header=T,sep="")
t12014<-read.csv("MMR1summary2014.txt",header=T,sep="")
t22014<-read.csv("MMR2summary2014.txt",header=T,sep="")

table_nir_mmr<-rbind(
t12006,
t22006,
t12007,
t22007,
t12008,
t22008,
t12009,
t22009,
t12010,
t22010,
t12011,
t22011,
t12012,
t22012,
t12013,
t22013,
t12014,
t22014)
colnames(table_nir_mmr)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
yr_nir<-c('MMR1 2006',
'MMR2 2006',
'MMR1 2007',
'MMR2 2007',
'MMR1 2008',
'MMR2 2008',
'MMR1 2009',
'MMR2 2009',
'MMR1 2010',
'MMR2 2010',
'MMR1 2011',
'MMR2 2011',
"MMR1 2012",
"MMR2 2012",
"MMR1 2013",
"MMR2 2013",
"MMR1 2014",
"MMR2 2014")
table_nir_mmr$Year <-yr_nir
table_nir_mmr<-table_nir_mmr[c(10,1:7)]
table_nir_mmr[2:7]<-table_nir_mmr[2:7]*100
latex(table_nir_mmr, file="", table.env=FALSE,rowname=NULL)

>>>>

\usepackage{Sweave}
\begin{document}
\input{draftfinalreport_v4-concordance}

@@

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand, 2013} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 17\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve to measles, it was estimated that approximately 54\% of those would require additional vaccination to ensure measles does not persist and outbreaks do not occur, leading to an additional 402,010 effective vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially immunised population) values often include or exceed one, including 2013--2014 outbreaks that were ongoing at the time of analysis. As analysed from data until 12 June 2014, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of the first 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost of those measles cases that attend hospital is estimated to be \$1,877 per attending case. Approximately 17\% of measles cases attend hospital.
\item The mean public health service cost per case (not including hospitalisation costs) is \$1,765.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccination-related costs required to exceed approximately \$1,117 per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs together.
\item The B/C ratio differed across DHBs with those with large populations and more na\"{\i}ve in absolute numbers most likely to benefit from immunisation campaigns. 
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 54\% of the currently 17\% na\"{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@
